Refined Flare Rates for Stars Younger than 250\,Myr

### 2023-09-26 src/tex/ms.tex
updated scripts

\section{Results}\label{sec:res}

\begin{figure*}[ht!]
    \script{mcmc_results.py}
    \begin{centering}
        \includegraphics[width=2\textwidth]{figures/mcmc_results.pdf}
        \caption{
            Measured flare-frequency distribution slopes, $\alpha$, as a function
            of stellar effective temperature, $T_\textrm{eff}$ and age.
        }
        \label{fig:mcmc_results}
    \end{centering}
\end{figure*}

\subsection{Flare Rates by Association}

\subsection{Flare Rates by Age}

\subsection{Flare Rates by Effective Temperature}

@@

>>>>

\subsection{Measuring Rotation Periods}

\cite{seligman22} demonstrated that stars with low Rossby numbers $R_0 < 0.13$ have
shallower flare frequency distribution slopes, indicative of more high energy flares
originating from these sources. In addition to understanding flare statistics across
young ages, we aim to expand this sample by measuring the rotation periods, $P_\textrm{rot}$,
for the stars in our sample. To do this, we used \texttt{michael}\footnote{\url{https://github.com/ojhall94/michael}},
an open-source Python package that robustly measures $P_\textrm{rot}$ using a combination
traditional Lomb-Scargle periodograms and wavelet transformations (Hall et al. submitted).
\texttt{michael} measures $P_\textrm{rot}$ using the \texttt{eleanor} package, which
extracts light curves from the TESS Full-Frame Images \citep[FFIs;][]{feinstein19}.
We ran \texttt{michael} on all stars from which flares were identified. The estimated
$P_\textrm{rot}$ were vetted by-eye, from the \texttt{michael} diagnostic plots. In
total, we robustly measured \textcolor{red}{n} $P_\textrm{rot}$ across our sample of
3,983 stars.

\section{Results}\label{sec:results}

We analyze our new flare sample from three perspectives. First, we perform the standard
FFD fitting of a power-law to the distribution of flare energies. Second, we fit the FFD with the
prescription in \cite{gershberg72}, which fits for both the FFD slope and y-intercept.
Third, we fit a truncated power-las to the distribution of flare amplitudes, to determine
if there is a correlation between $R_0$ and flare distributions.

The number of stars, and consequently flares, per each association varied greatly,
due to the limited number of stars observed at TESS 2-minute cadence. Therefore, instead
of measuring FFD properties as a function of association, we opted to group stars
by effective temperature, $T_\textrm{eff}$, and average adopted association age. We
grouped stars in the following $T_\textrm{eff}$ space: M-stars below the fully
convective boundary ($T_\textrm{eff} = 2300 - 3400$\,K), early type M-stars
($T_\textrm{eff} = 3400 - 3850$\,K), late K-stars ($T_\textrm{eff} = 3850 - 4440$\,K),
early K-stars ($T_\textrm{eff} = 4440 - 5270$\,K), and G-stars ($T_\textrm{eff} = 5270 - 5930$\,K).
We did not include any stars hotter than $T_\textrm{eff} > 5930$\,K, as these stars
are dominated by noise in the TESS observations. Additionally, we grouped stars
in the following age space: $4-10$\,Myr, $10-20$\,Myr, $20-40$\,Myr, $40-50$\,Myr,
$70-80$\,Myr, $120-150$\,Myr, and $150-300$\,Myr. We note that there is a gap in
age from $50-70$\,Myr, which could be expanded with the identification of more
associations in this age range. However, for the purposes of this work, we do not
include additional sources which may fall in this age range.

\subsection{Standard Power-Law Fits}\label{subsec:stats}

From the $T_\textrm{eff}$ and age bins described above, we binned the flares in each
subgroup and fit their FFD slopes, approximated as a power-law. Flares were binned
into 25 bins in log-space from $10^{27} - 10^{35}$\,erg. We fit the FFDs from the
energy bin with the maximum flare rate and energies higher than that. We opt to do
this as bins of lower energies may be incomplete, and the turnover in the FFD cannot
be accurately modeled as a power-law.

We fit the FFD using the MCMC method implemented in \texttt{emcee} \citep{goodman10, emcee}
and fit for the slope, $\alpha$, y-intercept, $b$, and an additional noise term, $f$, which
accounts for an underestimation of the errors on each bin. We initialized the MCMC fit with
300 walkers and ran our fit over 5000 steps. Upon visual inspection, we discarded the
first 100 steps; onwards the steps were fully burned-in. The full FFDs are presented
in Figure~\ref{fig:simple_ffd_all}, along with 100 samples from the MCMC fit. The measured
FFD slopes, $\alpha$ are presented in Figure~\ref{fig:mcmc_results}. We approximate the
error on the slope as the lower 16\textsuperscript{th} and upper 84\textsuperscript{th}
percentiles from the MCMC fit.

\begin{figure*}[ht!]
    \script{mcmc_results.py}
    \begin{centering}
        \includegraphics[width=3\textwidth]{figures/mcmc_results.pdf}
        \caption{
            Measured flare-frequency distribution slopes, $\alpha$, as a function
            of stellar effective temperature, $T_\textrm{eff}$ and age.
        }
        \label{fig:mcmc_results}
    \end{centering}
\end{figure*}

\subsection{Fitting for $\alpha$ and $\beta$}

\subsection{Truncated Power-Law Fits}

We follow the prescription presented in \cite{seligman22}. Namely, we fit a truncated
power-law distribution of the form

\begin{equation}
  dp/dA \propto A^{-\alpha_T} e^{-A/A_*}
\end{equation}

where $A$ is the amplitude of the flare, $A_*$ is a flare amplitude cutoff parameter
and $\alpha_T$ is the slope, rather than
$\alpha$. We fit the slopes using the MCMC method implemented in \texttt{emcee}
\citep{goodman10, emcee}, using the log-likelihood function in \cite{seligman22}.
We fit for $A_*$ and $\alpha_T$. We initialized the MCMC fit with textcolor{red}{n}
walkers and evaluated the fit over textcolor{red}{n} steps. The first textcolor{red}{n}
steps were discarded upon visual inspection. The results are presented in Figure~\ref{fig:trunc}.

\section{Discussion}\label{sec:discuss}

\subsection{Applications to ULLYSES}

Measuring Far-Ultraviolet (FUV) total flux output and flare rates are
essential for understanding the contributions of flares to exoplanet atmospheric
mass-loss and disequilibrium chemistry. An equivalent, but significantly smaller sample
to the stars studied here are the those observed as part of the \textit{Hubble Space Telescope}
(HST) Ultraviolet Legacy Library of Young Stars as Essential Standards (ULLYSES)
program. This survey observed 71 K- to M-type T Tauri stars in nine young
galactic associations.

We can extend the results presented here to estimate how many flares may be present
in these important FUV observations.

Additionally, we analyzed the TESS light curves for all of the ULLYSES targets, when
available at 2-minute cadence. We identified a total of \textcolor{red}{n} flares.
This is inconsistent with the overall flare statistics identified in this work, and is
overall unexpected for young stars which are known to be more magnetically active
\citep{ilin20, feinstein20b}. This could suggest that there is some bias in the way
the ULLYSES sample was constructed, and that it is unexpected to find any flares in the
FUV data. Alternatively, it could be that the ongoing accretion luminosity results
in a brighter host star, such that the flares we normally would see are drowned
out by the increased baseline luminosity. We do not include this flare in our
overall statistical analysis of TESS flare rates for young stars.

@@

This work made use of the open-source package, \textcolor{red}{\textit{showyourwork!}}
\citep{luger2021}, which promotes reproducible publications.

\appendix
\restartappendixnumbering

\section{Supplemental Material}\label{appendix:supp_ffds}

Miscellaneous figures and such that people might want but I don't need to show

\begin{figure*}[ht!]
    \script{simple_ffd_all.py}
    \begin{centering}
        \includegraphics[width=\textwidth]{figures/simple_ffd_all.pdf}
        \caption{
            Flare frequency distributions (FFDs) for subgroups of stars, clustered
            by age and effective temperature, $T_\textrm{eff}$. Flares were binned
            into 25 bins in log-space from $10^{27} - 10^{35}$\,erg. We fit the FFD
            from the turn-over in the binned flares, likely a result of very low-energy
            flares being missed in our flare-detection algorithm. The bins used to fit
            the FFD are shown in black, while all bins are shown in gray. We ran an MCMC
            fit to these distributions with a simple power law; 100 random samples from these
            fits are over-plotted in orange. We fit distributions with $> 3$ bins.
            The best-fit slopes from these fits are presented in Figure~\ref{fig:mcmc_results}.
        }
        \label{fig:simple_ffd_all}
    \end{centering}
\end{figure*}
